---
title: "Exploratory Data Analysis - Retail"
author: Dhriti Vohra
output: html_document
---
# DHRITI VOHRA

## *Data Science & Business Analytics Internship @ The Sparks Foundation*

**TASK #3**
**Exploratory Data Analysis - Retail**
**(Level - Beginner)**
**#GRIPMAY21**
   
* Perform ‘Exploratory Data Analysis’ on dataset ‘SampleSuperstore’
* As a business manager, try to find out the weak areas where you can
work to make more profit.
* What all business problems you can derive by exploring the data?
* You can choose any of the tool of your choice
(Python/R/Tableau/PowerBI/Excel/SAP/SAS)
* Dataset: <https://bit.ly/3i4rbWl>     

```{r}
#Importing the dataset in R:
retail <- read.csv("C:/Users/dhrit/Downloads/SampleSuperstore.csv")

#Taking a glimpse of dataset
head(retail)
```

**This is the data for a Super Store in US.**

```{r}
#Knowing the structure of the dataset
str(retail)
#Checking whether there is any null values in the dataset or not
sum(is.na(retail))
```

*There are 9994 Rows and 13 columns in the dataset.*

```{r}
#Get to know about the column and dimensions of your dataset
ncol(retail)
dim(retail)
```

```{r}
#Checking for unique values in the data
sapply(retail,function(x) length(unique(x)))
```


```{r}
#Comparing Sales and Profit
plot(density(retail$Profit),main="Overlaying Graphs",ylab="",type="l",col="blue",xlim=c(-500,3000),ylim=c(0,0.060))
polygon(density(retail$Sales),col="pink")
points(density(retail$Sales),col="red",type="l")
legend("topright",c("Profit","Sales"),fill = c("blue","red"))

```

*From the plot, we can observe that Profit is more than Sales but at some places Profit could be increased.*

```{r}
#To visualize the Category wise relationship between each quantitaive variable of the data, we draw scatter plots
library(ggplot2)
p1 <- ggplot(retail,aes(Postal.Code,fill=Category))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p2 <- ggplot(retail,aes(Sales,Postal.Code,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p3 <- ggplot(retail,aes(Quantity,Postal.Code,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p4 <- ggplot(retail,aes(Discount,Postal.Code,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p5 <- ggplot(retail,aes(Profit,Postal.Code,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p6 <- ggplot(retail,aes(Sales,fill=Category))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p7 <- ggplot(retail,aes(Postal.Code,Sales,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p8 <- ggplot(retail,aes(Quantity,Sales,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p9 <- ggplot(retail,aes(Discount,Sales,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p10 <- ggplot(retail,aes(Profit,Sales,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p11 <- ggplot(retail,aes(Quantity,fill=Category))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p12 <- ggplot(retail,aes(Postal.Code,Quantity,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p13 <- ggplot(retail,aes(Sales,Quantity,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p14 <- ggplot(retail,aes(Discount,Quantity,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p15 <- ggplot(retail,aes(Profit,Quantity,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p16 <- ggplot(retail,aes(Discount,fill=Category))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p17 <- ggplot(retail,aes(Postal.Code,Discount,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p18 <- ggplot(retail,aes(Sales,Discount,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p19 <- ggplot(retail,aes(Quantity,Discount,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p20 <- ggplot(retail,aes(Profit,Discount,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p21 <- ggplot(retail,aes(Profit,fill=Category))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p22 <- ggplot(retail,aes(Postal.Code,Profit,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p23 <- ggplot(retail,aes(Sales,Profit,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p24 <- ggplot(retail,aes(Quantity,Profit,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
p25 <- ggplot(retail,aes(Discount,Profit,fill=Category,col=Category))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
library("ggpubr")
annotate_figure(ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13,p14,p15,p16,p17,p18,p19,p20,p21,p22,p23,p24,p25,ncol=5,nrow = 5,common.legend = TRUE, legend = "top"),top = text_grob("Pairplot of each column based on Category",face="bold",size = 14))
```

```{r}
#To visualize the Region wise relationship between each quantitative variable of the data, we draw scatter plots
library("ggplot2")
r1 <- ggplot(retail,aes(Postal.Code,fill=Region))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r2 <- ggplot(retail,aes(Sales,Postal.Code,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r3 <- ggplot(retail,aes(Quantity,Postal.Code,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r4 <- ggplot(retail,aes(Discount,Postal.Code,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r5 <- ggplot(retail,aes(Profit,Postal.Code,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r6 <- ggplot(retail,aes(Sales,fill=Region))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r7 <- ggplot(retail,aes(Postal.Code,Sales,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r8 <- ggplot(retail,aes(Quantity,Sales,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r9 <- ggplot(retail,aes(Discount,Sales,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r10 <- ggplot(retail,aes(Profit,Sales,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r11 <- ggplot(retail,aes(Quantity,fill=Region))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r12 <- ggplot(retail,aes(Postal.Code,Quantity,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r13 <- ggplot(retail,aes(Sales,Quantity,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r14 <- ggplot(retail,aes(Discount,Quantity,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r15 <- ggplot(retail,aes(Profit,Quantity,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r16 <- ggplot(retail,aes(Discount,fill=Region))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r17 <- ggplot(retail,aes(Postal.Code,Discount,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r18 <- ggplot(retail,aes(Sales,Discount,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r19 <- ggplot(retail,aes(Quantity,Discount,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r20 <- ggplot(retail,aes(Profit,Discount,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r21 <- ggplot(retail,aes(Profit,fill=Region))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r22 <- ggplot(retail,aes(Postal.Code,Profit,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r23 <- ggplot(retail,aes(Sales,Profit,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r24 <- ggplot(retail,aes(Quantity,Profit,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
r25 <- ggplot(retail,aes(Discount,Profit,fill=Region,col=Region))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
library("ggpubr")
annotate_figure(ggarrange(r1,r2,r3,r4,r5,r6,r7,r8,r9,r10,r11,r12,r13,r14,r15,r16,r17,r18,r19,r20,21,r22,r23,r24,r25,ncol=5,nrow = 5,common.legend = TRUE, legend = "top"),top = text_grob("Pairplot of each column based on Region",face="bold",size = 14))
```

```{r}
#To visualize the Segment wise relationship between each quantitative variable of the data, we draw scatter plots
s1 <- ggplot(retail,aes(Postal.Code,fill=Segment))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s2 <- ggplot(retail,aes(Sales,Postal.Code,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s3 <- ggplot(retail,aes(Quantity,Postal.Code,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s4 <- ggplot(retail,aes(Discount,Postal.Code,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s5 <- ggplot(retail,aes(Profit,Postal.Code,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s6 <- ggplot(retail,aes(Sales,fill=Segment))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s7 <- ggplot(retail,aes(Postal.Code,Sales,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s8 <- ggplot(retail,aes(Quantity,Sales,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s9 <- ggplot(retail,aes(Discount,Sales,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s10 <- ggplot(retail,aes(Profit,Sales,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s11 <- ggplot(retail,aes(Quantity,fill=Segment))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s12 <- ggplot(retail,aes(Postal.Code,Quantity,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s13 <- ggplot(retail,aes(Sales,Quantity,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s14 <- ggplot(retail,aes(Discount,Quantity,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s15 <- ggplot(retail,aes(Profit,Quantity,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s16 <- ggplot(retail,aes(Discount,fill=Segment))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s17 <- ggplot(retail,aes(Postal.Code,Discount,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s18 <- ggplot(retail,aes(Sales,Discount,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s19 <- ggplot(retail,aes(Quantity,Discount,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s20 <- ggplot(retail,aes(Profit,Discount,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s21 <- ggplot(retail,aes(Profit,fill=Segment,col=Segment))+geom_density(alpha=0.3)+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s22 <- ggplot(retail,aes(Postal.Code,Profit,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s23 <- ggplot(retail,aes(Sales,Profit,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s24 <- ggplot(retail,aes(Quantity,Profit,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
s25 <- ggplot(retail,aes(Discount,Profit,fill=Segment,col=Segment))+geom_point()+theme(axis.text.x = element_blank(),axis.text.y = element_blank())
library("ggpubr")
annotate_figure(ggarrange(s1,s2,s3,s4,s5,s6,s7,s8,s9,s10,s11,s12,s13,s14,s15,s16,s17,s18,s19,s20,s21,s22,s23,s24,s25,ncol=5,nrow = 5,common.legend = TRUE, legend = "top"),top = text_grob("Pairplot of each column based on Segment",face="bold",size = 14))
```


```{r}
#To observe the correlation between Postal Code, Sales, Quantity, Discount and Profit, make a correlation matrix
retail_m <- cor(retail[c(FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,TRUE,TRUE,TRUE,TRUE)])
retail_m

```


```{r}
#Using "reshape" package to "melt" data 
library("reshape")
data_melt <- melt(retail_m)
#Plotting a Heatmap to visualize correlation among variables-Postal Code,      Sales,Quantity,Discount and Profit
library("ggplot2")
ggp <- ggplot(data_melt,aes(X1,X2))+geom_tile(aes(fill=value))+scale_fill_distiller(palette = "RdPu")+geom_text(aes(label=round(value,2)),nudge_x = 0.25, nudge_y = 0.25,check_overlap = T)+ggtitle("Heatmap")
ggp
```

*From the HEATMAP,we can observe that:* 

* Sales and Profit are moderately correlated.
* Sales and Quantity have LOW Positive correlation.
* Discount and Quantity are LEAST correlated to each other.
* Profit and Discount have NEGATIVE Correlation.

```{r}
#Count Plot for Quantitative Variables in the data
b1 <- ggplot(retail,aes(x=Category,fill=Category,col=Category))+geom_bar()+
  theme(axis.text.x =  element_text(angle = 45,hjust=1))
b2 <- ggplot(retail,aes(x=Segment,fill=Segment,col=Segment))+geom_bar()+
  theme(axis.text.x =  element_text(angle = 45,hjust=1))
b3 <- ggplot(retail,aes(x=Ship.Mode,fill=Ship.Mode,col=Ship.Mode))+geom_bar()+
  theme(axis.text.x =  element_text(angle = 45,hjust=1))
b4 <- ggplot(retail,aes(x=Region,fill=Region,col=Region))+geom_bar()+
  theme(axis.text.x =  element_text(angle = 45,hjust=1))
library("ggpubr")
ggarrange(b1,b2,b3,b4,ncol=2,nrow = 2)
```

*We observe that Furniture Category and West Region has large domination in Super Store Dealings, Consumer Segment contributes more to the Superstore and Standard Class Ship Mode is widely used for shipping the products.*

```{r}
#Bar plot for Sub Category
library("ggplot2")
ggplot(retail,aes(x=Sub.Category,fill=Sub.Category,col=Sub.Category))+geom_bar()+theme(axis.text.x =  element_text(angle = 90,hjust=1),legend.position = "none" )
```

```{r}
#Barplot for State
ggplot(retail,aes(x=State,fill=State,col=State))+geom_bar()+theme(axis.text.x =  element_text(angle = 90,hjust=1),legend.position = "none"  )
```

```{r}
#Bar Plot for Quantity
ggplot(retail,aes(x=Quantity,fill=Quantity,col=Quantity))+geom_bar(fill="pink")+theme(axis.text.x =  element_text(angle = 90,hjust=1),legend.position = "none"  )
```

```{r}
#Bar Plot for Discount
ggplot(retail,aes(x=Discount,fill=Discount,col=Discount))+geom_bar(fill="indianred3")+theme(axis.text.x =  element_text(angle = 90,hjust=1),legend.position = "none"  )
```

```{r}
#Looking at the Distribution of the data
par(mfrow=c(2,2))
colnames <- dimnames(retail)[[2]]
for(i in 10:13){
  d <- density(retail[,i])
  plot(d, type="n",main=colnames[i])
  polygon(d,col="cornflowerblue",border="grey")
}
```

```{r}
#Plotting and Analyzing State wise dealings 
ggplot(retail,aes(State))+geom_bar(fill="#0073C2FF")+labs(x="States",y="Frequency/Number of deals")+ggtitle("State Wise Dealings")+theme(axis.text.x =  element_text(size=9,angle = 45,hjust=1))
```

* *State with the HIGHEST Deals : California* 
* *State with the LOWEST Deals : Wyoming *

```{r}
retail_s <- data.frame(table(retail$State))
mean(retail_s$Freq)
```
*So, Average number of deals per state is 204*

```{r}
#Analyzing City wise dealings
retail_c <- data.frame(table(retail$City))
retail_c_order <- retail_c[order(retail_c$Freq,decreasing = TRUE),]
gg <- retail_c_order[1:50,]

ggplot(gg, aes(x=Var1, y=Freq)) + geom_bar(fill="#0073C2FF",stat = "identity")+labs(x="Cities",y="Frequency/Number of deals")+ggtitle("City Wise Dealings")+theme(axis.text.x =  element_text(size=9,angle = 45,hjust=1))
```

*City with the HIGHEST Deals : New York City* 

```{r}
mean(retail_c$Freq)
```

*Average number of deals per city is 19.*

```{r}
#Analyzing Segment wise Profit, Discount and Sales
summary(retail$Segment)
mean(table(retail$Segment,retail$Sales,retail$Quantity,retail$Discount))
library("dplyr")
seg <- retail %>% group_by(Segment) %>% 
summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
ungroup() %>% arrange(-M_Sales) %>% data.frame()
seg
```

```{r}
#Visualising Segment wise Sales, Discount and Profit
seg1 <- data.frame(sapply(seg[2:4],function(y) (y/(sum(y)))*100))
new_seg <- cbind(seg$Segment,seg1)
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))

pie1 <- ggplot(new_seg,aes(x="",y=M_Sales,fill=seg$Segment))+geom_bar(stat="identity",color="black",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(seg$Segment,"\n",round(M_Sales,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Sales")+ scale_fill_hue()

pie2 <- ggplot(new_seg,aes(x="",y=M_Discount,fill=seg$Segment))+geom_bar(stat="identity",color="black",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(seg$Segment,"\n",round(M_Discount,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Discount")+ scale_fill_hue()  

pie3 <- ggplot(new_seg,aes(x="",y=M_Profit,fill=seg$Segment))+geom_bar(stat="identity",color="black",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(seg$Segment,"\n",round(M_Profit),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Profit")+ scale_fill_hue()

library("ggpubr")
ggarrange(pie1,pie2,pie3,ncol=3,nrow = 1)
```

* Home Office contributes to the maximum number of Sales.
* Maximum Discount is offered by Consumer and Corporate Segments.
* Maximum profit earning segment is Home Office segment.

```{r}
summary(retail$State)
mean(table(retail$State,retail$Sales,retail$Quantity,retail$Discount))
```

```{r}
#Analyzing State wise Profit, Discount and Sales
library("dplyr")
sta <- retail %>% group_by(State) %>% 
summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
ungroup() %>% arrange(-M_Sales) %>% data.frame()
sta_o <- sta[order(sta$State),]
head(sta_o)
```

```{r}
#Visualizing State wise Sales, Discount and Profit
ggplot(sta_o,aes(x= State,y=M_Profit))+geom_bar(fill="#0073C2FF",stat = "identity")+theme(axis.text.x =  element_text(angle = 45,hjust=1),legend.position = "none" )+geom_text(aes(label = round(M_Profit,2)), col="black",fontface="bold",position = position_dodge(0.9),vjust=0.5,size=3)+labs(x = "State",y="Profit per state",title="State Wise Profit Analysis")
```

* Highest Profit earning State : Vermont   
* Lowest Profit earning State : Ohio

```{r}
sta1 <- data.frame(sapply(abs(sta_o[2]),function(y) (y/(sum(y)))*100))
sta1
new_sta <- cbind(sta_o$State,sta1)
new_sta
```


```{r}
#Visualizing State wise Sales
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))

ggplot(new_sta,aes(x="",y=M_Sales,fill=sta_o$State))+geom_bar(stat="identity",color="black",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank())+geom_text(aes(x=1.7,label = paste(round(M_Sales,1),"%")),position=position_stack(vjust=0.5),col="black",fontface="bold",size=3,angle=47)+labs(x = NULL, y = NULL, fill = NULL, title = "State wise Sales")+ scale_fill_hue()
```

* Highest amount of Sales : Wyoming (11.8%)
* Lowest amount of Sales : South Dakota (0.8%)

```{r}
#Visualizing State wise Discount
ggplot(sta_o[,c(1,3)], aes(x=State, y=M_Discount)) + geom_bar(fill="#0073C2FF",stat = "identity")+labs(x="State",y="Discount")+ggtitle("State Wise Analysis of Discount")+theme(axis.text.x =  element_text(size=8,angle = 45,hjust=1))
```

*State offering Highest Discount : Illinois*

```{r}
#Analyzing City wise Profit
library("dplyr")
cit <- retail %>% group_by(City) %>% 
summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
ungroup() %>% arrange(-M_Sales) %>% data.frame()

low_cit <- head(cit[cit$M_Profit<0,][,c(1,4)],30)
high_cit <- head(cit[cit$M_Profit>0,][,c(1,4)],30)
```

```{r}
#Cities earning LOW Profit OR Cities suffering LOSS
library("ggplot2")
ggplot(low_cit,aes(x=City,y=M_Profit))+geom_bar(fill="pink",stat = "identity")+labs(x="City",y="Profit")+ggtitle("City Wise analysis of Low Profit OR Loss")+theme(axis.text.x =  element_text(size=9,angle = 45,hjust=1))
```

*Bethlehem is the city earning lowest profit.*

```{r}
#Cities earning HIGH Profit
library("ggplot2")
ggplot(high_cit,aes(x=City,y=M_Profit))+geom_bar(fill="pink",stat = "identity")+labs(x="City",y="Profit")+ggtitle("City Wise analysis of Profit")+theme(axis.text.x =  element_text(size=9,angle = 45,hjust=1))
```

*Jamestown is the highest profit earning city.*

```{r}
#Analyzing Quantity wise Sales, Profit and Discount
library("dplyr")
quat <- retail %>% group_by(Quantity) %>% 
  summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
  ungroup() %>% arrange(-M_Sales) %>% data.frame()
qua <- quat[order(quat$Quantity),]
qua
quat1 <- data.frame(sapply(qua[2:4],function(y) (y/(sum(y)))*100))
new_quat <- cbind(qua$Quantity,quat1)
```

```{r}
#Visualising Quantity wise Sales, Discount and Profit
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))

pi1 <- ggplot(new_quat,aes(x="",y=M_Sales,fill=qua$Quantity))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank())+geom_text(aes(x=1.3,label = paste0(new_quat$Quantity,"\n",round(M_Sales,1),"%","")),position=position_stack(vjust=0.6),col="white")+labs(x = NULL, y = NULL, fill = NULL, title = "Sales")

pi2 <- ggplot(new_quat,aes(x="",y=M_Discount,fill=qua$Quantity))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank())+geom_text(aes(x=1.3,label = paste(new_quat$Quantity,"\n",round(M_Discount,1),"%","")),position=position_stack(vjust=0.5),col="white")+labs(x = NULL, y = NULL, fill = NULL, title = "Discount")

pi3 <- ggplot(new_quat,aes(x="",y=M_Profit,fill=qua$Quantity))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank())+geom_text(aes(x=1.3,label = paste(new_quat$Quantity,"\n",round(M_Profit),"%","")),position=position_stack(vjust=0.5),col="white")+labs(x = NULL, y = NULL, fill = NULL, title = "Profit")

library("ggpubr")
ggarrange(pi1,pi2,pi3,ncol=3,nrow = 1,common.legend = TRUE, legend = "bottom")
```

*Quantity number for high sales and profit is 13*

```{r}
#Analyzing Category wise Sales, Profit and Discount
library("dplyr")
cat <- retail %>% group_by(Category) %>% 
  summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
  ungroup() %>% arrange(-M_Sales) %>% data.frame()
cat
cat1 <- data.frame(sapply(cat[2:4],function(y) (y/(sum(y)))*100))
new_cat <- cbind(cat$Category,cat1)
```

```{r}
#Visualizing Category wise Sales, Profit and Discount
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))

pi_1 <- ggplot(new_cat,aes(x="",y=M_Sales,fill=cat$Category))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(cat$Category,"\n",round(M_Sales,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Sales")+ scale_fill_hue()

pi_2 <- ggplot(new_cat,aes(x="",y=M_Discount,fill=cat$Category))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(cat$Category,"\n",round(M_Discount,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Discount")+ scale_fill_hue()          

pi_3 <- ggplot(new_cat,aes(x="",y=M_Profit,fill=cat$Category))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(cat$Category,"\n",round(M_Profit),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Profit")+ scale_fill_hue()

library("ggpubr")
ggarrange(pi_1,pi_2,pi_3,ncol=3,nrow = 1)

```

* Highest Sales : Technology (49.1%)
* Highest Profit : Technology (73%)
* Minimum Sales : Office Supplies (12.9%)
* Minimum Profit : Furniture (8%)
* Maximum Discount : Furniture (37.5%)
* Minimum Discount : Technology (28.5%)

*Since, Sales and Profit are moderately correlated,High Sales of Technology leads to High Profit.*

```{r}
#Analyzing Sub-Category Wise Sales, Profit and Discount
library("dplyr")
sub_cat <- retail %>% group_by(Sub.Category) %>% 
  summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
  ungroup() %>% arrange(-M_Sales) %>% data.frame()
sub_cat
sub_cat1 <- data.frame(sapply(sub_cat[2:4],function(y) (y/(sum(y)))*100))
new_subcat <- cbind(sub_cat$Sub.Category,sub_cat1)

```

```{r}
#Visualizing Sub-Category Wise Sales
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))

ggplot(new_subcat,aes(x="",y=M_Sales,fill=sub_cat$Sub.Category))+geom_bar(stat="identity",color="black",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank())+geom_text(aes(x=1.5,label = paste0(round(M_Sales,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Sub-Category wise Sales")+ scale_fill_hue()

```

*Copiers have the HIGHEST Sales.*

```{r}
#Visualizing Sub-Category Wise Discount
ggplot(new_subcat,aes(x="",y=M_Discount,fill=sub_cat$Sub.Category))+geom_bar(stat="identity",color="black",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank())+geom_text(aes(x=1.3,label = paste0(round(M_Discount,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Sub- Category wise Discount")+ scale_fill_hue()          

```

*Binders, Machines and then Tables have High Discounts.*

```{r}
#Visualizing Sub-Category Wise Profit
library("tidyr")
sub_tall <- sub_cat %>% gather(key = sales_profit,value=Value,c(M_Sales,M_Profit))
sub_tall
ggplot(sub_tall,aes(x=Sub.Category,y=Value))+geom_bar(stat = "identity",aes(fill=sales_profit),position="dodge")+theme(axis.text.x =  element_text(size=9,angle = 45,hjust=1))+ggtitle("Comparing Sub-Category Wise Profit and Sales")
```

*Copiers have Highest Profit as well as Sales.*

```{r}
#Region Wise Analysis Of Profit, Sales and Discount
library("dplyr")
reg <- retail %>% group_by(Region) %>% 
  summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
  ungroup() %>% arrange(-M_Sales) %>% data.frame()
reg
reg1 <- data.frame(sapply(reg[2:4],function(y) (y/(sum(y)))*100))
new_reg <- cbind(reg$Region,reg1)
```

```{r}
#Visualising Region wise Sales, Profit and Discount
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))
pr_1 <- ggplot(new_reg,aes(x="",y=M_Sales,fill= reg$Region))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(new_reg$`reg$Region`,"\n",round(M_Sales,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Sales")+ scale_fill_hue()
pr_2 <- ggplot(new_reg,aes(x="",y=M_Discount,fill= reg$Region))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(new_reg$`reg$Region`,"\n",round(M_Discount,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Discount")+ scale_fill_hue()          
pr_3 <- ggplot(new_reg,aes(x="",y=M_Profit,fill= reg$Region))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(new_reg$`reg$Region`,"\n",round(M_Profit),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Profit")+ scale_fill_hue()

library("ggpubr")
ggarrange(pr_1,pr_2,pr_3,ncol=3,nrow = 1)
```

* South Region has Highest percentage of Sales (26.2%)
* Central Region offers Maximum Discount (37.4%), has Minimum sales (23.4%) and earns Minimum    Profit(15%)
* West Region has High amount of Profit (30%) and offers Minimum Discount (17%)

```{r}
summary(retail$Ship.Mode)
mean(table(retail$Ship.Mode,retail$Sales,retail$Quantity,retail$Discount))
```

```{r}
#Analyzing Ship Mode Wise Sales, Discount and Profit
library("dplyr")
ship <- retail %>% group_by(Ship.Mode) %>% 
  summarize(M_Sales=mean(Sales),M_Discount=mean(Discount),M_Profit=mean(Profit)) %>% 
  ungroup() %>% arrange(-M_Sales) %>% data.frame()
ship1 <- data.frame(sapply(ship[2:4],function(y) (y/(sum(y)))*100))
ship1
new_ship <- cbind(ship$Ship.Mode,ship1)
new_ship
```

```{r}
#Visualizing Ship Mode Wise Sales, Discount and Profit
library(ggplot2)
blank_theme <- theme_minimal()+theme(axis.title.x = element_blank(),axis.title.y = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.ticks = element_blank(),plot.title=element_text(size=14, face = "bold"))

psh_1 <- ggplot(new_ship,aes(x="",y=M_Sales,fill=ship$Ship.Mode))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(ship$Ship.Mode,"\n",round(M_Sales,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Sales")+ scale_fill_hue()

psh_2 <- ggplot(new_ship,aes(x="",y=M_Discount,fill=ship$Ship.Mode))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(ship$Ship.Mode,"\n",round(M_Discount,1),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Discount")+ scale_fill_hue()          

psh_3 <- ggplot(new_ship,aes(x="",y=M_Profit,fill=ship$Ship.Mode))+geom_bar(stat="identity",color="white",width = 1)+coord_polar("y", start=0) + blank_theme+ theme(axis.text.x=element_blank(),legend.position = 'FALSE')+geom_text(aes(label = paste0(ship$Ship.Mode,"\n",round(M_Profit),"%","")),position=position_stack(vjust=0.5),col="black")+labs(x = NULL, y = NULL, fill = NULL, title = "Profit")+ scale_fill_hue()

library("ggpubr")
ggarrange(psh_1,psh_2,psh_3,ncol=3,nrow = 1)
```

* Same Day Shipment has HIGH Sales (25.5%)
* Discounts (26.7%) and Profits (27%) are HIGH in First Class Shipping Mode 
* Standard class earns Minimum Profit (27%)
***

>Conclusions :

1. When comparing Profit and Sales,Profit is Higher than Sales but still it can be improved in some areas.
2. The State of California and Wyoming have Highest and Lowest Sales respectively.
3. New York City dominates in City Wise deals.
4. Segment Wise Analysis : 
    + Home office contributes to maximum Sales and Profit. 
    + Maximum Discounts are offered by Consumer and Corporate Segments and have less Profits.
5. Highest Profit earning State is Vermont and Lowest Profit earning State is Ohio.
6. Ohio,Colorado,North Carolina,Tennessee,Texas,Pennsylvania,Illinois are some States suffering from LOSS.
7. Illinois is the State offering Highest Discount.
8. Bethlehem and Jamestown are the lowest and Highest Profit earning cities respectively.
9. Category Wise Analysis : 
    + Sales and Profit are Highest in Technology (49.1%) while it offers Minimum Discount (28.5%)
    + Sales are minimum in Office Supplies (12.9%)
    + Profit is Minimum in Furniture (8%) and it is offering Maximum Discount (37.5%)
    + *Sales and Profit are moderately correlated,High Sales of Technology leads to High Profit.*
10. Sub Category Analysis : 
     + Copiers have Highest Profit as well as Sales.
     + Binders, Machines and Tables offer very High Discounts.
     + Tables are suffering from loss.
11. Region wise Analysis : 
     + Central Region offers Maximum Discount (37.4%), has Minimum sales (23.4%) and earns Minimum Profit(15%)
     + West Region has High amount of Profit (30%) and offers Minimum Discount (17%)
12. Ship Mode wise Analysis : 
   + Discounts (26.7%) and Profits (27%) are HIGH in First Class Shipping Mode 
   + Standard class earns Minimum Profit (27%)
                              
**Area of Improvement :**

*Since, Profit and Discount have NEGATIVE Correlation i.e., if High Discount is offered then, most probably it will lead to LOW profits. This happens in many cases here like -*`

* Maximum Discounts are offered by Consumer and Corporate Segments and have less Profits.
* Illinois is the State offering Highest Discount and is suffering from loss.
* Sales and Profit are Highest in Technology (49.1%) was it offers Minimum Discount (28.5%).
* Profit is Minimum in Furniture (8%) as it is offering Maximum Discount (37.5%).
* Tables offer very High Discounts. and are suffering from loss.
* Central Region offers Maximum Discount (37.4%) and earns Minimum Profit(15%) while West        Region has High amount of Profit (30%) and offers Minimum Discount (17%).

***By observing the pattern it is advisable for Central Region to offer less discounts so that they could earn considerable amount of profit. Same is the case with the states like Illinois, Ohio,Texas which offers large discounts.***




